Statistical analysis script

### 2023-04-05 code/analysis_code/statistical_analysis.qmd
add vip

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
library(ranger)
library(knitr)
library(skimr)
```
## Load Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
skim(training_data_final)
```
## First Health Outcome of Interest: Illnesses

@@

# Plotting predictions vs Illnesses for Linear model
lm_rs1%>%
  collect_predictions()%>%
  filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with main predictor")
ggsave("Predicted Illnesses VS Illnesses by linear regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
lm_rs2%>%
  collect_predictions()%>%
  #filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with all predictors")
ggsave("Predicted Illnesses VS Illnesses by linear regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")
# Plotting predictions vs Illnesses for Poisson model
poisson_rs1%>%
  collect_predictions()%>%
  #filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with main predictor")
ggsave("Predicted Illnesses VS Illnesses by poisson regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
poisson_rs2%>%
  collect_predictions()%>%
  #filter(Illnesses<500)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with all predictors")
ggsave("Predicted Illnesses VS Illnesses by poisson regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")

@@

### All predictors
poisson_rs4 = poisson_wf4%>%
  fit_resamples(resamples=hosp_fold,
    control=control_resamples(save_pred=TRUE))
```
### Model Evaluation
```{r}
# Checking metrics for Linear Regression Model
lm_rs3%>%
  collect_metrics()
lm_rs4%>%
  collect_metrics()
# Checking metrics for poisson regression model
poisson_rs3%>%
  collect_metrics()
poisson_rs4%>%
  collect_metrics()
# Checking Parameters for Linear Regression Model
lm_fit3=lm_wf3%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

lm_fit4=lm_wf4%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

# Checking Parameters for Poisson Regression Model
poisson_fit3=poisson_wf3%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

poisson_fit4=poisson_wf4%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

# Plotting predictions vs Hospitalizations for Linear model
lm_rs3%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with main predictor",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
lm_rs4%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with all predictors",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")
# Plotting predictions vs Hospitalizations for Poisson model
poisson_rs3%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with main predictor",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
poisson_rs4%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with all predictors",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")

@@

## Machine learning approach
## Random Forest
### Predicting First Outcome `Illnesses` using all predictors
```{r}
# Model Specification
rf_spec=rand_forest(mtry=tune(),min_n=tune(),trees=1000)%>%
  set_engine("ranger")%>%
  set_mode("regression")
illness_all_rec=recipe(Illnesses~.,data=illness_data)%>%
  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
  step_dummy(all_nominal())
# Creating workflow
rf_wf1=workflow()%>%
  add_recipe(illness_all_rec)%>%
  add_model(rf_spec)
# Training and tuning model using space filling design
set.seed(345)
doParallel::registerDoParallel()
rf_tune1=rf_wf1%>%
  tune_grid(
    resamples=illness_fold,
    grid=15,
    control=control_grid(save_pred=TRUE),
    metrics = metric_set(rmse)
  )
```
### Model evaluation
```{r}
rf_tune1%>%
  autoplot()
rf_best1=rf_tune1%>%
  select_best()
rf_final1=rf_wf1%>%
  finalize_workflow(rf_best1)
rf_fit1=rf_final1%>%
  fit(illness_data)
rf_predict1=rf_fit1%>%
  predict(illness_data)
illness_data%>%
  select(Illnesses)%>%
  bind_cols(rf_predict1)%>%
  ggplot(aes(Illnesses,.pred))+
  geom_point()+
  geom_abline(intercept = 0,slope=1)+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted illnesses VS illnesses by random forest model")
ggsave("Predicted illnesses VS illnesses by random forest model.png",path=images_stats_path,width=3000,height=2000,units="px")
```

>>>>

# Data Handling and Modeling
library(tidyverse)
library(tidymodels)
library(poissonreg)
library(ranger)
library(knitr)
library(skimr)
library(vip)
library(caret)
```
## Load Data
```{r}
# Path to Data
data_location <- here(
  "data","processed_data","processed_training_data.rds")
# Import Data
training_data_final <- readRDS(data_location)
skim(training_data_final)
```
## First Health Outcome of Interest: Illnesses

@@

# Checking variable importance for Linear and Poisson regression models
lm_wf1%>%
  fit(illness_data)%>%
  extract_fit_engine()%>%
  vip()
lm_wf2%>%
  fit(illness_data)%>%
  extract_fit_engine()%>%
  vip()
poisson_wf1%>%
  fit(illness_data)%>%
  extract_fit_engine()%>%
  vip()
poisson_wf2%>%
  fit(illness_data)%>%
  extract_fit_engine()%>%
  vip()
# Plotting predictions vs Illnesses for Linear model
lm_rs1%>%
  collect_predictions()%>%
  #filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with main predictor")
ggsave("Predicted Illnesses VS Illnesses by linear regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
lm_rs2%>%
  collect_predictions()%>%
  #filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with all predictors")
ggsave("Predicted Illnesses VS Illnesses by linear regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")
# Plotting predictions vs Illnesses for Poisson model
poisson_rs1%>%
  collect_predictions()%>%
  #filter(Illnesses<300)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with main predictor")
ggsave("Predicted Illnesses VS Illnesses by poisson regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
poisson_rs2%>%
  collect_predictions()%>%
  #filter(Illnesses<500)%>%
  ggplot(aes(Illnesses,.pred,color=id2,group=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Illnesses",title="Predicted Illnesses VS Illnesses by linear regression with all predictors")
ggsave("Predicted Illnesses VS Illnesses by poisson regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")

@@

### All predictors
poisson_rs4 = poisson_wf4%>%
  fit_resamples(resamples=hosp_fold,
    control=control_resamples(save_pred=TRUE))
```
### Model Evaluation
```{r}
# Checking metrics for Linear Regression Model
lm_rs3%>%
  collect_metrics()
lm_rs4%>%
  collect_metrics()
# Checking metrics for poisson regression model
poisson_rs3%>%
  collect_metrics()
poisson_rs4%>%
  collect_metrics()
# Checking Parameters for Linear Regression Model
lm_wf3%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

lm_wf4%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

# Checking Parameters for Poisson Regression Model
poisson_wf3%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

poisson_wf4%>%
  fit(hosp_data)%>%
  tidy()%>%
  arrange(-estimate)

# Checking variable importance for linear and Poisson regression models
lm_wf3%>%
  fit(hosp_data)%>%
  extract_fit_engine()%>%
  vip()
lm_wf4%>%
  fit(hosp_data)%>%
  extract_fit_engine()%>%
  vip()
poisson_wf3%>%
  fit(hosp_data)%>%
  extract_fit_engine()%>%
  vip()
poisson_wf4%>%
  fit(hosp_data)%>%
  extract_fit_engine()%>%
  vip()

# Plotting predictions vs Hospitalizations for Linear model
lm_rs3%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with main predictor",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
lm_rs4%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with all predictors",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by linear regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")
# Plotting predictions vs Hospitalizations for Poisson model
poisson_rs3%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with main predictor",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with main predictor.png",path=images_stats_path,width=3000,height=2000,units="px")
poisson_rs4%>%
  collect_predictions()%>%
  #filter(Hospitalizations<50)%>%
  ggplot(aes(Hospitalizations,.pred,color=id2))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_abline(intercept=0,slope=1)+
  labs(y="Predicted Hospitalizations",title="Predicted Hospitalizations VS Hospitalizations \nby linear regression with all predictors",color="Fold")
ggsave("Predicted Hospitalizations VS Hospitalizations by poisson regression with all predictors.png",path=images_stats_path,width=3000,height=2000,units="px")

@@

## Machine learning approach
## Random Forest
### Predicting First Outcome `Illnesses` using all predictors
```{r}
# Model Specification
rf_spec=rand_forest(mtry=tune(),min_n=tune(),trees=1000)%>%
  set_engine("ranger",importance = 'impurity')%>%
  set_mode("regression")
illness_all_rec=recipe(Illnesses~.,data=illness_data)%>%
  step_other(State,Location,`IFSAC Category`,Simplified_Etiology,threshold = 0.01)%>%
  step_dummy(all_nominal())
# Creating workflow
rf_wf1=workflow()%>%
  add_recipe(illness_all_rec)%>%
  add_model(rf_spec)
# Training and tuning model using space filling design
set.seed(345)
doParallel::registerDoParallel()
rf_tune1=rf_wf1%>%
  tune_grid(
    resamples=illness_fold,
    grid=15,
    control=control_grid(save_pred=TRUE),
    metrics = metric_set(rmse)
  )
```
### Model evaluation
```{r}
rf_tune1%>%
  autoplot()
rf_best1=rf_tune1%>%
  select_best()
rf_final1=rf_wf1%>%
  finalize_workflow(rf_best1)
rf_fit1=rf_final1%>%
  fit(illness_data)
# Plotting predicted illness VS Illnesses
rf_predict1=rf_fit1%>%
  predict(illness_data)
illness_data%>%
  select(Illnesses)%>%
  bind_cols(rf_predict1)%>%
  ggplot(aes(Illnesses,.pred))+
  geom_point()+
  geom_abline(intercept = 0,slope=1)+
  geom_smooth(se=FALSE)+
  labs(y="Predicted Illnesses",title="Predicted illnesses VS illnesses by random forest model")
ggsave("Predicted illnesses VS illnesses by random forest model.png",path=images_stats_path,width=3000,height=2000,units="px")
```
